---
title: "Homework 1"
output: github_document
---
Name: Yuchen Gu Uni:yg2964

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library (ggplot2)
library(dplyr)
```

# Problem 1

## Section 1 - Data import and Description

```{r}

data("penguins", package = "palmerpenguins")

```

The penguins dataset contains information about three penguins species : Adelie, Chinstrap, and Gentoo. It has `r nrow(penguins)` observations and `r ncol(penguins)` variables.

This dataset includes details about each penguin such as its species, the island where it lives, and various body measurements. These measurements include bill length and depth, flipper length, and body mass. It also records the sex of each penguin and the year of observation. From this data, we can see that the average flipper length of the penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm, and their average body mass is `r round(mean(penguins$body_mass_g, na.rm = TRUE), 2)` g.

It's important to note that the dataset isn't perfect. There are some missing values: `r sum(is.na(penguins$body_mass_g))` for body mass and `r sum(is.na(penguins$sex))` for sex.

Scientists can use this data to study differences between penguin species, how penguins might vary across islands, and differences between male and female penguins. This helps in understanding penguin biology and their adaptation to the Antarctic environment. 

## Section 2 Scatterplot Creation
```{r echo=FALSE, message=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(na.rm = TRUE)
ggsave("penguins_scatterplot.png", width = 8, height = 5)
```

This scatterplot effectively illustrates the relationship between bill length and flipper length across three penguin species. Distinct clusters are visible, with Gentoo penguins (blue) showing longer flippers and bills, Adelie penguins (red) having shorter measurements in both dimensions, and Chinstrap penguins (green) occupying a middle ground. A positive correlation between bill length and flipper length is apparent within each species. 





